---
title: "Lab4"
subtitle: "Exercises"
author: "Irimie Fabio"
output: 
  pdf_document:
    toc: true
    keep_tex: true 
---

# Exercise 1

## A
Plot the Probability Mass Function for the Binomial distribution with 
$n=18$ and $p=\frac{1}{3}$.
Calculate:

  1. $P(X=3)$
```{r, collapse=TRUE}
dbinom(3, 18, 1 / 3)
```
  2. $P(X \ge 3)$
```{r, collapse=TRUE}
1 - pbinom(2, 18, 1 / 3)
# or
pbinom(2, 18, 1 / 3, lower.tail = FALSE)
```
  3. $P(1 \le X > 5)$
```{r, collapse=TRUE}
pbinom(4, 18, 1 / 3) - pbinom(0, 18, 1 / 3)
```
  4. $P(X \ge 15)$
```{r, collapse=TRUE}
pbinom(14, 18, 1 / 3, lower.tail = FALSE)
```

```{r}
library(ggplot2)
df <- data.frame(x = 0:18, y = dbinom(0:18, 18, 1 / 3))

ggplot(df, aes(x = x, y = y, fill = "PMF")) +
  geom_col() +
  scale_x_continuous(breaks = 0:18) +
  scale_fill_manual(values = "red") +
  labs(
    title = "Binomial Distribution",
    x = "x",
    y = "P(X = x)"
  )
```

## B
Plot the Cumulative Distribution Function for the Poisson distribution 
with $\lambda = 3$. Calculate:

  1. $P(X=3)$
```{r, collapse=TRUE}
lambda <- 3
dpois(3, lambda)
```
  2. $P(X \ge 3)$
```{r, collapse=TRUE}
ppois(2, lambda, lower.tail = FALSE)
```
  3. $P(1 \le X > 5)$
```{r, collapse=TRUE}
ppois(4, lambda) - ppois(0, lambda)
```
  4. $P(X \ge 15)$
```{r, collapse=TRUE}
ppois(14, lambda, lower.tail = FALSE)
```

```{r}
df <- data.frame(x = 0:18, y = ppois(0:18, lambda))

ggplot(df, aes(x = x, y = y, color = "CDF")) +
  geom_step() +
  scale_x_continuous(breaks = 0:18) +
  scale_color_manual(values = "blue") +
  labs(
    title = "Poisson Distribution",
    x = "x",
    y = "P(X <= x)"
  )
```

# Exercise 2
Demonstrate that a Poisson r.v. may be used as an approximation for a
binomial r.v.

## A
```{r}
n <- c(20, 30, 40, 100)
p <- c(1 / 4, 1 / 6, 1 / 8, 1 / 20)

pmf <- matrix(NA, nrow = 21, ncol = 4)

for (i in 1:4) {
  pmf[, i] <- dbinom(0:20, n[i], p[i])
}

pmf <- as.data.frame(pmf)

colnames(pmf) <- paste("Binomial", n, round(p, 2), sep = "_")

pmf$Poisson <- dpois(0:20, n * p)

pmf$X <- 0:20
```
## B
```{r}
library(reshape2)

df_plot <- melt(pmf, id.vars = "X")
df_plot

library(ggplot2)
library(RColorBrewer) # Color palettes

ggplot(df_plot, aes(x = X, y = value, color = variable)) +
  geom_line() +
  scale_color_manual(values = c(brewer.pal(4, "PRGn"), "red")) +
  labs(
    title = "Binomial vs Poisson PMF",
    x = "x",
    y = "P(X = x)"
  )
```

# Exercise 3

## A
Generate N=1000 random numbers from a binomial distribution with n=9
trials and p=0.8. Thus each of the 1000 random numbers will be an
integer between 0 and 9.
```{r}
set.seed(123)
n <- 9
p <- 0.8
t <- 1000

valori <- rbinom(t, n, p)
```

## B
Plot the experimental probability using the geom_bar() function.
```{r}
df <- data.frame(x = valori)

ggplot(df, aes(x = valori)) +
  geom_bar() +
  scale_x_continuous(breaks = 0:9) +
  labs(
    title = "Experimental Probability",
    x = "Successes",
    y = "Frequency"
  )
```

## C
For each value of the x-axis obtained in the previous plot, compute the real
probability mass function and add it in the plot as red dots using geom_point().
```{r}
p_teo <- data.frame(p = dbinom(0:9, n, p))

ggplot(df, aes(x = valori)) +
  geom_bar(aes(y = after_stat(prop))) +
  geom_point(
    data = p_teo,
    aes(
      x = 0:9, y = p,
      color = "red"
    ),
  ) +
  scale_x_continuous(breaks = 0:9) +
  labs(
    title = "Experimental vs Real Probability",
    x = "Successes",
    y = "Frequency"
  )
```
