---
title: "Lab4"
subtitle: "Exercises"
author: "Irimie Fabio"
output: 
  pdf_document:
    toc: true
    keep_tex: true 
---

# Exercise 1

## A
Create the Lab3 project. Use the same structure used for Lab1 and Lab2:
scripts, plots and data directories.

## B
Write a function to calculate the sum of integer numbers from 1 to n
```{r}
sum_integer <- function(n) {
  sum <- 0
  for (i in 1:n) {
    sum <- sum + i
  }
  return(sum)
}

cat("The sum of the first 10 integers is: ", sum_integer(10), "\n")
```

## C
Write a function to calculate the product of integers from 1 to n, also
known as n!
```{r}
prod_integer <- function(n) {
  val <- n
  for (i in (n - 1):1) {
    val <- val * i
  }
  return(val)
}

cat("The factorial of 5 is: ", prod_integer(5), "\n")
```

## D
Try C. but do it recursively (hint: call the function itself inside the
loop, remember to return 1 when n is equal to 0)
```{r}
factorial <- function(n) {
  if (n == 0) {
    return(1)
  } else {
    val <- n * factorial(n - 1)
  }

  return(val)
}

cat("The factorial of 5 is: ", factorial(5), "\n")
```

# Exercise 2

## A
Simulate the tossing of a fair dice and verify through the definition
that the event $E = \{ 2,3 \}$ has probability $\frac{1}{3}$.
$S = \{ 1,2,3,4,5,6 \}$; $E = \{ 2,3 \}$; $P(E) = \frac{1}{3}$

(hint: generate a sequence of integer random numbers between 1 and 6
using the sample() function)
```{r}
library(ggplot2)

n <- 100000 # Number of experiments
e <- c(2, 3) # Event of interest

# Outcomes of interest
set.seed(123)
res <- sample(x = c(1:6), size = n, replace = TRUE)

# Outcomes of E (1 when in E, 0 otherwise)
ne <- ifelse(res %in% e, 1, 0)
```
## B
Plot the first 40 outcomes of the experiment.
```{r}
ggplot(
  data = data.frame(x = 1:40, y = res[1:40]),
  aes(
    x = x,
    y = y,
    color = factor(res[1:40])
  )
) +
  geom_point() +
  scale_color_manual(values = rainbow(6)) +
  labs(
    title = "First 40 outcomes of the dice",
    x = "Number of experiments", y = "Value"
  )
```

## C
Plot the convergence of $P(E)$ at the value obtained from the classical
definition $\frac{1}{3}$. (hint: the frequentist approach says that, 
as the number of trials approaches infinity,
the relative frequency will converge exactly to the true probability)
```{r}
# Probability of E
pe <- sum(ne) / n
cat("The probability of E is: ", pe, "\n")

cum_pe <- cumsum(ne) / 1:n

df <- data.frame(x = 1:n, y = cum_pe)
ggplot(data = df, aes(
  x = x, y = y,
)) +
  geom_line() +
  geom_hline(yintercept = 1 / 3, col = "red") +
  scale_x_continuous(trans = "log10") +
  labs(
    title = "Cumulative probability of E",
    x = "Number of experiments", y = "Cumulative probability"
  )
```

# Exercise 3

## A
Simulate the tossing of a fair dice and consider the following events:
$A=\{1,2\};\;B=\{2,3,6\};\;C=\{1,4,5\}$. (hint: compute $P(A),P(B),P(C)$).
```{r, collapse=TRUE}
n <- 100000 # Number of experiments

a <- c(1, 2) # Event A
b <- c(2, 3, 6) # Event B
c <- c(1, 4, 5) # Event C

res <- sample(x = c(1:6), size = n, replace = TRUE)

pa <- sum(res %in% a) / n
pa
pb <- sum(res %in% b) / n
pb
pc <- sum(res %in% c) / n
pc
```

## B
Verify that $A$ and $B$ are independent and that $B$ and $C$ are 
dependent.
```{r}
# A and B
pab <- sum(res %in% a & res %in% b) / n

# B and C
pbc <- sum(res %in% b & res %in% c) / n

cat("A and B are independent: ", pab == pa * pb, "\n")
cat("B and C are dependent: ", pbc != pb * pc, "\n")
```

# Exercise 4

## A
Generate a sequence of $N=10000$ random numbers that simulate the
throwing of a dice.
```{r}
n <- 10000
set.seed(123)
res <- sample(x = c(1:6), size = n, replace = TRUE)
```

## B
Then simulate the throwing of a second dice.
```{r}
res2 <- sample(x = c(1:6), size = n, replace = TRUE)
```

## C
Plot the absolute and relative frequencies for A. and the relative
frequency for the sum of the two dice for point B. using the geom_bar()
or geom_col() functions.
```{r}
df1 <- data.frame(x = 1:6, y = table(res) / n)
df2 <- data.frame(x = 2:12, y = table(res + res2) / n)

ggplot(data = df1, aes(
  x = factor(x),
  y = y.Freq
)) +
  geom_col() +
  labs(
    title = "Absolute frequency of the dice",
    x = "Value", y = "Absolute frequency"
  ) +
  theme_minimal()
```
```{r}
ggplot(data = df2, aes(
  x = factor(x),
  y = y.Freq
)) +
  geom_col() +
  labs(
    title = "Relative frequency of the sum of the dice",
    x = "Value", y = "Relative frequency"
  ) +
  theme_minimal()
```

# Exercise 5

## A
Four people are in a room. What is the probability that no two of them
celebrate their birthday on the same day of the year?
```{r}
people <- 4
days <- 365

n_bdays <- choose(days, people) * factorial(people)
p_bdays <- n_bdays / days^people
p_bdays
```

## B
$n$ people are in a room. What is the probability that no two of them
celebrate their birthday on the same day of the year? Try this with
$n$ from 1 to 100 and plot the probability for each value of $n$.
```{r}
n <- 100
p_bdays <- rep(NA, n)

for (i in 1:n) {
  p_bdays[i] <- choose(days, i) * factorial(i) / days^i
}

df <- data.frame(x = 1:n, y = p_bdays)
ggplot(data = df, aes(
  x = x, y = y
)) +
  geom_line() +
  labs(
    title = "Probability of no two people having the same birthday",
    x = "Number of people", y = "Probability"
  ) +
  theme_minimal()
```

# Exercise 6
A laboratory blood test is 99% effective in detecting a certain 
disease when it is, in fact, present. However, the test also yields 
a “false positive” (FP – Type I error) result for 1% of the healthy 
persons tested. If 0.5% of the population actually has the disease,
what is the probability a person has the disease given that his test
result is positive? (Hint: Use D as the event "the tested person has 
the disease" and E as "The result of the test is positive").
```{r}
p_d <- 0.005
p_e_d <- 0.99
p_e_nd <- 0.01

p_d_e <- p_d * p_e_d / (p_d * p_e_d + (1 - p_d) * p_e_nd)
p_d_e
```
