---
title: "Exercise 2"
output: pdf
---

# Exercise 2
## A
Create 2 vectos twice: using the colon oprator and the seq(function)

vec1: 1, 2, 3, 4, 5, 6, 7, 8, 9, 10

vec2: 2, 7, 12

```{r}
vec1 <- 1:10
vec2 <- seq(2, 12, 5)
cat("Vec1: ", vec1, "\nVec2: ", vec2)
```

## B
Create a 4x2 matrix of all zeros and store it in a variable (mymat).
Then, replace the second row in the matrix with a vector consisting of
a 3 and a 6.
```{r}
mymat <- matrix(
  0,
  4,
  2
)
cat("Starting matrix:\n")
mymat
mymat[2, 1] <- 3
mymat[2, 2] <- 6
cat("Replaced matrix:\n")
mymat
```

## C
Create a vector x which consists of 20 equally spaced points in the range
from $-\pi$ to $+\pi$. Create a y vector which is sin(x).
```{r}
x <- seq(-pi, pi, length.out = 20)
x

y <- sin(x)
y
```

## D
Create a 4x6 matrix of random integers, each in the range 
from -5 to 5; store it in a variable (mat).
Create another matrix that stores for each element the 
absolute value of the corresponding element in
the original matrix (mat_pos).
```{r}
mat <- matrix(
  round(runif(4 * 6, -5, 5), digits = 0),
  4,
  6
)
mat_pos <- abs(mat)

mat
mat_pos
```

