---
title: "Lab1"
subtitle: "Exercise 3"
author: "Irimie Fabio"
output: 
  pdf_document:
    toc: true
    keep_tex: true 
---

## A
Load the sunspot.year dataset from the datasets package. Use
data("sunspot.year") and then sunspot.year to load it in the workspace.
```{r}
data("sunspot.year")
sunspot.year
```

## B
See the documentation to obtain information about the dataset and create
a sequence vector corresponding to the years. Call this variable year.
```{r}
data <- sunspot.year
year <- seq(start(data)[1], end(data)[1], 1)
year
```

## C
Create a variable called sunspot, containing the values from the dataset
```{r}
sunspot <- c(data)
sunspot
```

## D
Put together the variables into a data.frame object.
```{r}
x <- data.frame(
  Year = year,
  Sunspots = sunspot
)
x
```

## E
Make a line plot of sunspots vs. year.

```{r, echo=FALSE}
plot(
  year,
  sunspot,
  type = "l",
  xlab = "Year",
  ylab = "Sunspots"
)
```

## F
Superimpose data points as red asterisks. Add a second layer to the plot
by using the points() function. Use pch = "*" and col = "red" in the
points() arguments. 
```{r}
plot(
  year,
  sunspot,
  type = "l",
  xlab = "Year",
  ylab = "Sunspots"
)
points(year, sunspot, pch = "*", col = "red")
```

## G
Create a title ‘Sunspots by year’.
```{r}
plot(
  year,
  sunspot,
  type = "l",
  xlab = "Year",
  ylab = "Sunspots",
  main = "Sunspots by year"
)
points(year, sunspot, pch = "*", col = "red")
```

## H
Make a column with 3 panels for the plot created in G., a barplot of
sunspots (you can use the as.vector() function to convert a a data type 
to a vector data type), and a histogram of sunspots.
```{r}
par(mfrow = c(3, 1))

plot(
  year,
  sunspot,
  type = "l",
  xlab = "Year",
  ylab = "Sunspots",
  main = "Sunspots by year"
)
points(year, sunspot, pch = "*", col = "red")

barplot(
  year,
  sunspot,
  xlab = "Year",
  ylab = "Sunspots",
  main = "Sunspots by year"
)

plot(
  year,
  sunspot,
  type = "h",
  xlab = "Year",
  ylab = "Sunspots",
  main = "Sunspots by year"
)
```
## I
Save the plot in the ./plots directory of the project as a .png file.
```{r, warning=FALSE,message=FALSE}
png("./plots/sunspots.png")
plot(
  year,
  sunspot,
  type = "l",
  xlab = "Year",
  ylab = "Sunspots",
  main = "Sunspots by year"
)
points(year, sunspot, pch = "*", col = "red")
dev.off()
```

## J
Save the data frame as a .csv file in the ./data directory of the 
project.
```{r}
write.csv(x, "./data/data.csv")
```
