---
title: "Lab 1"
subtitle: "Exercise 2"
author: "Irimie Fabio"
output: 
  pdf_document:
    toc: TRUE
    keep_tex: TRUE 
---

## A
Create 2 vectos twice: using the colon oprator and the seq(function)

vec1: 1, 2, 3, 4, 5, 6, 7, 8, 9, 10

vec2: 2, 7, 12

```{r, collapse=TRUE}
vec1 <- 1:10
cat("Vec1.1: ", vec1, "\n")
vec1 <- seq(1, 10, 1)
cat("Vec1.2: ", vec1, "\n")

vec2 <- (1:12)[-c(1, 3:6, 8:11)]
cat("Vec2.1: ", vec2, "\n")
vec2 <- seq(2, 12, 5)
cat("Vec2.2: ", vec2, "\n")
```

## B
Create a 4x2 matrix of all zeros and store it in a variable (mymat).
Then, replace the second row in the matrix with a vector consisting of
a 3 and a 6.
```{r, collapse=TRUE}
mymat <- matrix(
  0,
  4,
  2
)
cat("Starting matrix:\n")
mymat
mymat[2, ] <- c(3, 6)
cat("Replaced matrix:\n")
mymat
```

## C
Create a vector x which consists of 20 equally spaced points in the range
from $-\pi$ to $+\pi$. Create a y vector which is sin(x).
```{r, collapse=TRUE}
x <- seq(-pi, pi, length.out = 20)
x

y <- sin(x)
y
```

## D
Create a 4x6 matrix of random integers, each in the range 
from -5 to 5; store it in a variable (mat).
Create another matrix that stores for each element the 
absolute value of the corresponding element in
the original matrix (mat_pos).
```{r, collapse=TRUE}
mat <- matrix(
  round(runif(4 * 6, -5, 5), digits = 0),
  4,
  6
)
mat_pos <- abs(mat)

mat
mat_pos
```

## E
Plot exp(x) for values of x ranging from -2 to 2 in steps of 0.1. Put 
an appropriate title on the plot, and label the axes.
<!-- echo = FALSE to remove code and leave only the plot -->
```{r, collapse=TRUE} 
x <- seq(-2, 2, 0.1)
y <- exp(x)
plot(x, y,
  type = "l",
  xlab = "x",
  ylab = "y",
  main = "Exponential function"
)
```

## F
Create a vector x with values ranging from 1 to 100 in steps of 5.
Create a vector y which is the square root of each value in x.
Plot these points. Now, use the barplot() function instead of plot()
to get a bar chart. Keep both plots together.
```{r, collapse=TRUE}
x <- seq(1, 100, 5)
y <- sqrt(x)
par(mfrow = c(1, 2))
plot(x, y)
barplot(x, y)
```
