---
title: "Lab6"
subtitle: "Exercises"
author: "Irimie Fabio"
output: 
  pdf_document:
    toc: true
    keep_tex: true 
---

# Exercise 1

## A
Compare the PDF's and the CDF's of the following Uniform continuous
random variables (hint: see page 3 of the slides).

Calculate:
1. $X1 \sim U(0,1)$
```{r}
a <- 0
b <- 1
seq <- seq(a - 0.5, b + 0.5, 0.01)
x1 <- data.frame(
  x = seq,
  pdf = dunif(seq, a, b),
  cdf = punif(seq, a, b)
)
```
2. $X2 \sim U(-3,2)$
```{r}
a <- -3
b <- 2
seq <- seq(a - 0.5, b + 0.5, 0.01)
x2 <- data.frame(
  x = seq,
  pdf = dunif(seq, a, b),
  cdf = punif(seq, a, b)
)
```
3. $X3 \sim U(2,4)$
```{r}
a <- 2
b <- 4
seq <- seq(a - 0.5, b + 0.5, 0.01)
x3 <- data.frame(
  x = seq,
  pdf = dunif(seq, a, b),
  cdf = punif(seq, a, b)
)
```
4. $X4 \sim U(0.8,2.5)$
```{r}
a <- 0.8
b <- 2.5
seq <- seq(a - 0.5, b + 0.5, 0.01)
x4 <- data.frame(
  x = seq,
  pdf = dunif(seq, a, b),
  cdf = punif(seq, a, b)
)
```

## B
Create a new figure with two vertical panels: first row PDFs, second row
CDFs (hint: use the library cowplot and the plot_grid() function to
arrange multiple ggplot objects).
```{r}
library(ggplot2)
library(cowplot)

pdfs <- ggplot() +
  geom_line(data = x1, aes(x = x, y = pdf), color = "red") +
  geom_line(data = x2, aes(x = x, y = pdf), color = "orange") +
  geom_line(data = x3, aes(x = x, y = pdf), color = "blue") +
  geom_line(data = x4, aes(x = x, y = pdf), color = "purple") +
  labs(title = "PDFs", x = "Values", y = "PDF") +
  theme_minimal()


cdfs <- ggplot() +
  geom_line(data = x1, aes(x = x, y = cdf), color = "red") +
  geom_line(data = x2, aes(x = x, y = cdf), color = "orange") +
  geom_line(data = x3, aes(x = x, y = cdf), color = "blue") +
  geom_line(data = x4, aes(x = x, y = cdf), color = "purple") +
  labs(title = "CDFs", x = "Values", y = "CDF") +
  theme_minimal()

plot_grid(pdfs, cdfs, ncol = 1)
```

## C
Determine the mean and variance of each random variable.
```{r}
x1_mean <- mean(runif(1000000, 0, 1))
x1_var <- var(runif(1000000, 0, 1))

x2_mean <- mean(runif(1000000, -3, 2))
x2_var <- var(runif(1000000, -3, 2))

x3_mean <- mean(runif(1000000, 2, 4))
x3_var <- var(runif(1000000, 2, 4))

x4_mean <- mean(runif(1000000, 0.8, 2.5))
x4_var <- var(runif(1000000, 0.8, 2.5))

table <- data.frame(
  x = c("X1", "X2", "X3", "X4"),
  mean = c(x1_mean, x2_mean, x3_mean, x4_mean),
  var = c(x1_var, x2_var, x3_var, x4_var)
)
table
```

# Exercise 2

## A
Determine and plot the values of the PDF for a normal distribution
with mean 3 and sd 2 for $x$ values between -1 and 7.
```{r}
mu <- 3
sd <- 2
seq <- seq(-1, 7, 0.01)
normal <- data.frame(
  x = seq,
  pdf = dnorm(seq, mu, sd)
)

ggplot(normal, aes(x = x, y = pdf)) +
  geom_line(aes(y = pdf, color = "red")) +
  labs(
    title = "PDF",
    subtitle = "mean = 3, sd = 2",
    x = "Values",
    y = "PDF"
  ) +
  theme_minimal()
```

## B
Determine and plot the corresponding values of the CDF
```{r}
mu <- 3
sd <- 2
seq <- seq(-1, 7, 0.01)
normal <- data.frame(
  x = seq,
  pdf = pnorm(seq, mu, sd)
)

ggplot(normal, aes(x = x, y = pdf)) +
  geom_line(aes(y = pdf, color = "red")) +
  labs(
    title = "PDF",
    subtitle = "mean = 3, sd = 2",
    x = "Values",
    y = "PDF"
  ) +
  theme_minimal()
```

## C
Determine for what values of $x$ the value of the CDF equals 0.025 
and 0.975 (hint: remember all the prefixes: p, d, r, q. See Lab 4,
slide 2).
```{r}
qnorm(0.025, mu, sd)
qnorm(0.975, mu, sd)
```

# Exercise 3

## A
Create a set of values ranging from xmin to xmax.
```{r}
mu <- 100 # the mean
sigma <- 15 # the standard deviation
xmin <- 70 # minumum x value for pdf and cdf plots
xmax <- 130 # maximum x value for pdf and cdf plots
n <- 100 # number of points of pdf and cdf plots
k <- 10000 # number of random draws (for histogram)

seq <- seq(xmin, xmax, length = n)
seq
```

## B
Draw k random numbers from a $N(\mu, \sigma^2)$ distribution.
```{r}
x <- rnorm(k, mu, sigma)

ggplot(data.frame(x = x), aes(x = x)) +
  geom_line(aes(y = ..density..), stat = "density") +
  labs(
    title = "Histogram",
    subtitle = "mean = 100, sd = 15",
    x = "Values",
    y = "Frequency"
  ) +
  theme_minimal()
```
## C
Create a new figure with 3 panels: PDF, CDF, and histogram of 
random values with 20 bins.
```{r}
pdf <- data.frame(
  x = seq,
  pdf = dnorm(seq, mu, sigma)
)

cdf <- data.frame(
  x = seq,
  cdf = pnorm(seq, mu, sigma)
)

histogram <- data.frame(
  x = x
)

pdf_plot <- ggplot(pdf, aes(x = x, y = pdf)) +
  geom_line(aes(y = pdf)) +
  labs(
    title = "PDF",
    subtitle = "mean = 100, sd = 15",
    x = "Values",
    y = "PDF"
  ) +
  theme_minimal()

cdf_plot <- ggplot(cdf, aes(x = x, y = cdf)) +
  geom_line(aes(y = cdf)) +
  labs(
    title = "CDF",
    subtitle = "mean = 100, sd = 15",
    x = "Values",
    y = "CDF"
  ) +
  theme_minimal()

histogram_plot <- ggplot(histogram, aes(x = x)) +
  geom_histogram(bins = 20) +
  labs(
    title = "Histogram",
    subtitle = "mean = 100, sd = 15",
    x = "Values",
    y = "Frequency"
  ) +
  theme_minimal()

plot_grid(pdf_plot, cdf_plot, histogram_plot, nrow = 1)
```

# Exercise 4
The number of visits per day to a website follows a Poisson
distribution with a mean of 120 visits per day. The duration of
each visit follows an exponential distribution with a mean time of
5 minutes. The percentage of visitors who purchase something is 
20%, and the bills follow a normal distribution with a mean of 
50€ and a standard deviation of 10€.

## A
Calculate the probability that at least 120 visits are made in 
a day.
```{r}
lambda <- 120
ppois(119, lambda, lower.tail = FALSE)
```

## B
Calculate the expected value and variance of the average duration
of a visit.
```{r}
lambda <- 1 / 5
mean <- 1 / lambda
var <- 1 / lambda^2
mean
var
```

## C
Draw the PDF for the random variable "duration of a visit".
```{r}
seq <- seq(0, 30, 0.01)
exponential <- data.frame(
  x = seq,
  pdf = dexp(seq, lambda)
)

ggplot(exponential, aes(x = x, y = pdf)) +
  geom_line(aes(y = pdf, color = "red")) +
  labs(
    title = "PDF",
    subtitle = "mean = 5",
    x = "Time (minutes)",
    y = "PDF"
  ) +
  theme_minimal()
```

## D
Calculate the probability that a visit lasts from 5 to 10 minutes.
```{r}
pexp(10, lambda) - pexp(5, lambda)
```

## E
Given that exactly 100 visits were made in a day, what is the 
probability to observe at least 20 purchases?
```{r}
lambda <- 0.2 * 100
ppois(19, lambda, lower.tail = FALSE)
```

## F
Compute the probabilities to observe a purchase between 40 and 60, 30 
and 70, 20 and 80.
```{r}
mu <- 50
sigma <- 10
pnorm(60, mu, sigma) - pnorm(40, mu, sigma)
pnorm(70, mu, sigma) - pnorm(30, mu, sigma)
pnorm(80, mu, sigma) - pnorm(20, mu, sigma)
```

## G
Simulate 1000 average days, compute the average daily revenue and 
compare it to the expected revenue.
```{r}
n <- 1000
lambda_visits <- 120
lambda_duration <- 1 / 5
purchases <- 0.2
mu_bills <- 50
sigma_bills <- 10

visits <- rpois(n, lambda_visits)
durations <- rexp(n, lambda_duration)
purchases <- rbinom(n, visits, purchases)
bills <- rnorm(n, mu_bills, sigma_bills)
revenue <- purchases * bills

daily_revenue <- mean(revenue)
expected_revenue <- lambda_visits * purchases * mu_bills

daily_revenue
mean(expected_revenue)
```
